---
title: "Gijon Air Pollution"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , echo = FALSE, warning= FALSE, message= FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(openair) # http://davidcarslaw.github.io/openair/
library(purrr)
library(lubridate)
library(ggplot2)
library(stringr)
library(knitr)
library(xts)
library(zoo)
library(gridExtra)
library(astsa)
library(rvest)
library(fpp2)
library(ranger)
library(broom)
library(RcppRoll)

```


### Preface
  
  This is the final project of a Master on Data Science. The main objective of this project is to show I can apply in a practical way the adquired knowledge during the master. 
  This is not a scientific paper. And I am not an specialist on air pollution. If you are interested in learning about air pollution my advice is to start visiting any of this web sites:
  
    - epa https://www.epa.gov/clean-air-act-overview
    - who https://www.who.int/airpollution/en/
    - eu https://www.eea.europa.eu/themes/air
    
    - Modelización de la contaminación por partículas PM10 en la aglomeración de Gijón ftp://ftp.asturias.es/asturias/medioambiente/Planes%20de%20mejora%20de%20la%20calidad%20del%20aire/Planes%202014-16/Estudios%20planes%202014-16/02-Modelizacion_contaminacion_PM10_Area_Gijon.pdf
    - Plan de mejora de la calidad del aire en la aglomeración área de Gijón (ES0309) https://www.asturias.es/medioambiente/articulos/ficheros/PLAN%20GIJON%20COMPLETO.pdf
    - ESTUDIO DE CONTRIBUCIÓN DE FUENTES EN LAS PARTÍCULAS PM10 EN SUSPENSIÓN EN LA AGLOMERACIÓN ÁREA DE GIJÓN (ES0309) Y EN LA
ZONA DE AVILÉS (ES0307)[here](ftp://ftp.asturias.es/asturias/medioambiente/Planes%20de%20mejora%20de%20la%20calidad%20del%20aire/Planes%202017-19/Estudios%20planes%202017-19/2.%20Estudio%20contribuci%F3n%20fuentes%20en%20part%EDculas%20PM10%20-%20Maqua.pdf)
    - Calidad del Aire y Salud en Asturias. Informe Epidemiológico 2016 https://www.astursalud.es/documents/31867/36150/Calidad+del+Aire+y+Salud.pdf/c389ab51-032c-245d-8cb8-505556521cb7
    
    Si estas interesado en aprender sobre este tema yo empezaria por echar un vistazo a la informacion que ofrece la EPA (), la OMS () y la EU (). Y si llegaste aqui porque quieres saber algo mas especifico sobre Gijon mi recomendacion es que leas los informes del Principado de Asturias...
  

  
### 1. Introduction

  The quality of the air is becoming a source of public concern. 
  - Construct a web dashboard to visualize the evolution of the main air pollution indicators of Gijon city. (Asturias; Spain).
  - Build a model to predict in the short term (<24h) the evolution of air pollutants concentration by monitoring station.


### 2. The data

I gathered the data used in this project from the open data web portal of the town hall of Gijón https://transparencia.gijon.es/. The data can be downloaded from  [here](https://transparencia.gijon.es/search/risp_dataset/page/1808-catalogo-de-datos?utf8=%E2%9C%93&search=aire+&search_sector=&search_format=&commit=Buscar&authenticity_token=j8%2F3CvCuPcDkrRe%2F1NR5RBp0t%2FOOosiA7724w3T2mB4%3D): 

I downloaded 18 csv files with air pollution and weather data of Gijón from years 2000 to 2017. I saved them in the "data" folder.
I downloaded two more files from this web, a csv file with the description of the variables and another csv file with information about the measurement stations. 



### 3. Methodology

### 4. Conclusions

### 5. References



