---
title: "Gijon Air Pollution"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , echo = FALSE, warning= FALSE, message= FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(openair) # http://davidcarslaw.github.io/openair/
library(purrr)
library(lubridate)
library(ggplot2)
library(stringr)
library(knitr)
library(xts)
library(zoo)
library(gridExtra)
library(astsa)
library(rvest)
library(fpp2)
library(ranger)
library(broom)
library(RcppRoll)

```


### Preface
  
  This is the final project of a Master on Data Science. 
  This is not a scientific paper. And I am not an specialist on air pollution. 
  The main objective of this project is to show I  

  
### 1. Introduction

  The quality of the air is becoming a source of public concern. 
  - Construct a web dashboard to visualize the evolution of the main air pollution indicators of Gijon city. (Asturias; Spain).
  - Build a model to predict in the short term (<24h) the evolution of air pollutants concentration by monitoring station.


### 2. The data

I gathered the data used in this project from the open data web portal of the town hall of Gijón https://transparencia.gijon.es/. The data can be downloaded from  [here](https://transparencia.gijon.es/search/risp_dataset/page/1808-catalogo-de-datos?utf8=%E2%9C%93&search=aire+&search_sector=&search_format=&commit=Buscar&authenticity_token=j8%2F3CvCuPcDkrRe%2F1NR5RBp0t%2FOOosiA7724w3T2mB4%3D): 

I downloaded 18 csv files with air pollution and weather data of Gijón from years 2000 to 2017. I saved them in the "data" folder.
I downloaded two more files from this web, a csv file with the description of the variables and another csv file with information about the measurement stations. 



### 3. Methodology

### 4. Conclusions

### 5. References



