---
title: "Train / test datasets for Python Models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading packages
```{r , warning= FALSE, message= FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(stringr)
library(knitr)
library(gridExtra)
library(ranger)
library(broom)
library(RcppRoll)
```

As we are going to start to work with auto regressive models, we are going to need dataframes with the values of the variable to forecast (the target) and lagged versions of itself as inputs. 

## PM10 datasets

First we create the initial datasets for training, testing and validation for the PM10 variable. 
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

# We load the constitucion_data dataset
constitucion_data <- readRDS("data_rds/constitucion_data.rds") 

# We extract the periods of data that we are going to use 

train_PM10_201401_201612 <- constitucion_data %>% filter(date_time_utc >= '2014-01-01 00:00:00',
                                       date_time_utc <= '2016-12-31 23:00:00') 

# We generate a smaller training dataset with the last three months of 2016
train_PM10_201610_201612 <- constitucion_data %>% filter(date_time_utc >= '2016-10-01 00:00:00',
                                       date_time_utc <= '2016-12-31 23:00:00') 
                                       

test_PM10_201701_201709 <- constitucion_data %>% filter(date_time_utc >= '2017-01-01 00:00:00',
                                       date_time_utc <= '2017-09-30 23:00:00') 

# We generate a smaller testing dataset with the first two weeks of 2016. 
test_PM10_20170101_20170114 <- constitucion_data %>% filter(date_time_utc >= '2017-01-01 00:00:00',
                                       date_time_utc <= '2017-01-14 23:00:00') 

validation_PM10_201710_201712 <- constitucion_data %>% filter(date_time_utc >= '2017-10-01 00:00:00',
                                       date_time_utc <= '2017-12-31 23:00:00') 

# We include all the datasets in a list
datasets_list <- list(train_PM10_201401_201612, train_PM10_201610_201612, test_PM10_201701_201709, test_PM10_20170101_20170114, validation_PM10_201710_201712)

# And we apply the same transformations to every dataframe in the list
datasets_list <- datasets_list %>% 
            map(~ select(., PM10)) %>% # We select the PM10 variable.
            map(~ mutate(., PM10 = replace_na(PM10, mean(PM10, na.rm = TRUE)))) %>% # We replace the NAs by the mean.
            map(~ rename(., PM10_0 = PM10)) # And we change the name of the variable to "PM10_0"


```

We add to the dataframes the PM10 lagged variables.
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

datasets_list <- datasets_list %>%
                  map(~ mutate(., PM10_1 = lag(PM10_0, 1),
                  PM10_2 = lag(PM10_0, 2),
                  PM10_3 = lag(PM10_0, 3),
                  PM10_4 = lag(PM10_0, 4),
                  PM10_5 = lag(PM10_0, 5),
                  PM10_6 = lag(PM10_0, 6),
                  PM10_7 = lag(PM10_0, 7),
                  PM10_8 = lag(PM10_0, 8),
                  PM10_9 = lag(PM10_0, 9),
                  PM10_10 = lag(PM10_0, 10),
                  PM10_11 = lag(PM10_0, 11),
                  PM10_12 = lag(PM10_0, 12),
                  PM10_13 = lag(PM10_0, 13),
                  PM10_14 = lag(PM10_0, 14),
                  PM10_15 = lag(PM10_0, 15),
                  PM10_16 = lag(PM10_0, 16),
                  PM10_17 = lag(PM10_0, 17),
                  PM10_18 = lag(PM10_0, 18),
                  PM10_19 = lag(PM10_0, 19),
                  PM10_20 = lag(PM10_0, 20),
                  PM10_21 = lag(PM10_0, 21),
                  PM10_22 = lag(PM10_0, 22),
                  PM10_23 = lag(PM10_0, 23),
                  PM10_24 = lag(PM10_0, 24),
                  PM10_25 = lag(PM10_0, 25),
                  PM10_26 = lag(PM10_0, 26),
                  PM10_27 = lag(PM10_0, 27),
                  PM10_48 = lag(PM10_0, 48),
                  PM10_72 = lag(PM10_0, 72),
                  PM10_96 = lag(PM10_0, 96),
                  PM10_120 = lag(PM10_0, 120),
                  PM10_144 = lag(PM10_0, 144),
                  PM10_168 = lag(PM10_0, 168))) %>%
                na.omit()


# We unlist the dataframes and we rename them again.
for (i in seq(datasets_list))
  assign(paste0("df", i), datasets_list[[i]])

train_PM10_201401_201612 <- df1
train_PM10_201610_201612 <- df2 
test_PM10_201701_201709 <- df3
test_PM10_20170101_20170114 <- df4 
validation_PM10_201710_201712 <- df5

```


We create and save all the y, X datasets (as csv and rds files)
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

y_train_PM10_201401_201612 <- train_PM10_201401_201612 %>% select(PM10_0)
X_train_PM10_201401_201612 <- train_PM10_201401_201612 %>% select(-PM10_0)

write_csv(X_train_PM10_201401_201612, "train_test/Prediction_Models_ML_0_AR/X_train_PM10_201401_201612.csv")
write_csv(y_train_PM10_201401_201612, "train_test/Prediction_Models_ML_0_AR/y_train_PM10_201401_201612.csv")

saveRDS(X_train_PM10_201401_201612, "train_test/Prediction_Models_ML_0_AR/X_train_PM10_201401_201612.rds")
saveRDS(y_train_PM10_201401_201612, "train_test/Prediction_Models_ML_0_AR/y_train_PM10_201401_201612.rds")


y_train_PM10_201610_201612 <- train_PM10_201610_201612 %>% select(PM10_0)
X_train_PM10_201610_201612 <- train_PM10_201610_201612 %>% select(-PM10_0)

write_csv(X_train_PM10_201610_201612, "train_test/Prediction_Models_ML_0_AR/X_train_PM10_201610_201612.csv")
write_csv(y_train_PM10_201610_201612, "train_test/Prediction_Models_ML_0_AR/y_train_PM10_201610_201612.csv")

saveRDS(X_train_PM10_201610_201612, "train_test/Prediction_Models_ML_0_AR/X_train_PM10_201610_201612.rds")
saveRDS(y_train_PM10_201610_201612, "train_test/Prediction_Models_ML_0_AR/y_train_PM10_201610_201612.rds")

y_test_PM10_20170101_20170114 <- test_PM10_20170101_20170114 %>% select(PM10_0)
X_test_PM10_20170101_20170114 <- test_PM10_20170101_20170114 %>% select(-PM10_0)

write_csv(X_test_PM10_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/X_test_PM10_20170101_20170114.csv")
write_csv(y_test_PM10_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/y_test_PM10_20170101_20170114.csv")


saveRDS(X_test_PM10_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/X_test_PM10_20170101_20170114.rds")
saveRDS(y_test_PM10_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/y_test_PM10_20170101_20170114.rds")

y_test_PM10_201701_201709 <- test_PM10_201701_201709 %>% select(PM10_0)
X_test_PM10_201701_201709 <- test_PM10_201701_201709 %>% select(-PM10_0)

write_csv(X_test_PM10_201701_201709, "train_test/Prediction_Models_ML_0_AR/X_test_PM10_201701_201709.csv")
write_csv(y_test_PM10_201701_201709, "train_test/Prediction_Models_ML_0_AR/y_test_PM10_201701_201709.csv")

saveRDS(X_test_PM10_201701_201709, "train_test/Prediction_Models_ML_0_AR/X_test_PM10_201701_201709.rds")
saveRDS(y_test_PM10_201701_201709, "train_test/Prediction_Models_ML_0_AR/y_test_PM10_201701_201709.rds")

y_validation_PM10_201710_201712 <- validation_PM10_201710_201712 %>% select(PM10_0)
X_validation_PM10_201710_201712 <- validation_PM10_201710_201712 %>% select(-PM10_0)

write_csv(X_validation_PM10_201710_201712, "train_test/Prediction_Models_ML_0_AR/X_validation_PM10_201710_201712.csv")
write_csv(y_validation_PM10_201710_201712, "train_test/Prediction_Models_ML_0_AR/y_validation_PM10_201710_201712.csv")

saveRDS(X_validation_PM10_201710_201712, "train_test/Prediction_Models_ML_0_AR/X_validation_PM10_201710_201712.rds")
saveRDS(y_validation_PM10_201710_201712, "train_test/Prediction_Models_ML_0_AR/y_validation_PM10_201710_201712.rds")

# We remove all objects from the workspace

rm(list = ls())

```


## NO2 datasets

And we generate the datasets for the NO2 pollutant. 
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

# We load the constitucion_data dataset
constitucion_data <- readRDS("data_rds/constitucion_data.rds") 

# We extract the periods of data that we are going to use 

train_NO2_201401_201612 <- constitucion_data %>% filter(date_time_utc >= '2014-01-01 00:00:00',
                                       date_time_utc <= '2016-12-31 23:00:00') 

# We generate a smaller training dataset with the last three months of 2016
train_NO2_201610_201612 <- constitucion_data %>% filter(date_time_utc >= '2016-10-01 00:00:00',
                                       date_time_utc <= '2016-12-31 23:00:00') 
                                       

test_NO2_201701_201709 <- constitucion_data %>% filter(date_time_utc >= '2017-01-01 00:00:00',
                                       date_time_utc <= '2017-09-30 23:00:00') 

# We generate a smaller testing dataset with the first two weeks of 2016. 
test_NO2_20170101_20170114 <- constitucion_data %>% filter(date_time_utc >= '2017-01-01 00:00:00',
                                       date_time_utc <= '2017-01-14 23:00:00') 

validation_NO2_201710_201712 <- constitucion_data %>% filter(date_time_utc >= '2017-10-01 00:00:00',
                                       date_time_utc <= '2017-12-31 23:00:00') 

# We include all the datasets in a list
datasets_list <- list(train_NO2_201401_201612, train_NO2_201610_201612, test_NO2_201701_201709, test_NO2_20170101_20170114, validation_NO2_201710_201712)

# And we apply the same transformations to every dataframe in the list
datasets_list <- datasets_list %>% 
            map(~ select(., NO2)) %>% # We select the NO2 variable.
            map(~ mutate(., NO2 = replace_na(NO2, mean(NO2, na.rm = TRUE)))) %>% # We replace the NAs by the mean.
            map(~ rename(., NO2_0 = NO2)) # And we change the name of the variable to "NO2_0"


```

We add to the dataframes the NO2 lagged variables.
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

datasets_list <- datasets_list %>%
                  map(~ mutate(., NO2_1 = lag(NO2_0, 1),
                  NO2_2 = lag(NO2_0, 2),
                  NO2_3 = lag(NO2_0, 3),
                  NO2_4 = lag(NO2_0, 4),
                  NO2_5 = lag(NO2_0, 5),
                  NO2_6 = lag(NO2_0, 6),
                  NO2_7 = lag(NO2_0, 7),
                  NO2_8 = lag(NO2_0, 8),
                  NO2_9 = lag(NO2_0, 9),
                  NO2_10 = lag(NO2_0, 10),
                  NO2_11 = lag(NO2_0, 11),
                  NO2_12 = lag(NO2_0, 12),
                  NO2_13 = lag(NO2_0, 13),
                  NO2_14 = lag(NO2_0, 14),
                  NO2_15 = lag(NO2_0, 15),
                  NO2_16 = lag(NO2_0, 16),
                  NO2_17 = lag(NO2_0, 17),
                  NO2_18 = lag(NO2_0, 18),
                  NO2_19 = lag(NO2_0, 19),
                  NO2_20 = lag(NO2_0, 20),
                  NO2_21 = lag(NO2_0, 21),
                  NO2_22 = lag(NO2_0, 22),
                  NO2_23 = lag(NO2_0, 23),
                  NO2_24 = lag(NO2_0, 24),
                  NO2_25 = lag(NO2_0, 25),
                  NO2_26 = lag(NO2_0, 26),
                  NO2_27 = lag(NO2_0, 27),
                  NO2_48 = lag(NO2_0, 48),
                  NO2_72 = lag(NO2_0, 72),
                  NO2_96 = lag(NO2_0, 96),
                  NO2_120 = lag(NO2_0, 120),
                  NO2_144 = lag(NO2_0, 144),
                  NO2_168 = lag(NO2_0, 168))) %>%
                na.omit()


# We unlist the dataframes and we rename them again.
for (i in seq(datasets_list))
  assign(paste0("df", i), datasets_list[[i]])

train_NO2_201401_201612 <- df1
train_NO2_201610_201612 <- df2 
test_NO2_201701_201709 <- df3
test_NO2_20170101_20170114 <- df4 
validation_NO2_201710_201712 <- df5

```


We create and save all the y, X datasets (as csv and rds files)
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

y_train_NO2_201401_201612 <- train_NO2_201401_201612 %>% select(NO2_0)
X_train_NO2_201401_201612 <- train_NO2_201401_201612 %>% select(-NO2_0)

write_csv(X_train_NO2_201401_201612, "train_test/Prediction_Models_ML_0_AR/X_train_NO2_201401_201612.csv")
write_csv(y_train_NO2_201401_201612, "train_test/Prediction_Models_ML_0_AR/y_train_NO2_201401_201612.csv")

saveRDS(X_train_NO2_201401_201612, "train_test/Prediction_Models_ML_0_AR/X_train_NO2_201401_201612.rds")
saveRDS(y_train_NO2_201401_201612, "train_test/Prediction_Models_ML_0_AR/y_train_NO2_201401_201612.rds")


y_train_NO2_201610_201612 <- train_NO2_201610_201612 %>% select(NO2_0)
X_train_NO2_201610_201612 <- train_NO2_201610_201612 %>% select(-NO2_0)

write_csv(X_train_NO2_201610_201612, "train_test/Prediction_Models_ML_0_AR/X_train_NO2_201610_201612.csv")
write_csv(y_train_NO2_201610_201612, "train_test/Prediction_Models_ML_0_AR/y_train_NO2_201610_201612.csv")

saveRDS(X_train_NO2_201610_201612, "train_test/Prediction_Models_ML_0_AR/X_train_NO2_201610_201612.rds")
saveRDS(y_train_NO2_201610_201612, "train_test/Prediction_Models_ML_0_AR/y_train_NO2_201610_201612.rds")

y_test_NO2_20170101_20170114 <- test_NO2_20170101_20170114 %>% select(NO2_0)
X_test_NO2_20170101_20170114 <- test_NO2_20170101_20170114 %>% select(-NO2_0)

write_csv(X_test_NO2_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/X_test_NO2_20170101_20170114.csv")
write_csv(y_test_NO2_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/y_test_NO2_20170101_20170114.csv")


saveRDS(X_test_NO2_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/X_test_NO2_20170101_20170114.rds")
saveRDS(y_test_NO2_20170101_20170114, "train_test/Prediction_Models_ML_0_AR/y_test_NO2_20170101_20170114.rds")

y_test_NO2_201701_201709 <- test_NO2_201701_201709 %>% select(NO2_0)
X_test_NO2_201701_201709 <- test_NO2_201701_201709 %>% select(-NO2_0)

write_csv(X_test_NO2_201701_201709, "train_test/Prediction_Models_ML_0_AR/X_test_NO2_201701_201709.csv")
write_csv(y_test_NO2_201701_201709, "train_test/Prediction_Models_ML_0_AR/y_test_NO2_201701_201709.csv")

saveRDS(X_test_NO2_201701_201709, "train_test/Prediction_Models_ML_0_AR/X_test_NO2_201701_201709.rds")
saveRDS(y_test_NO2_201701_201709, "train_test/Prediction_Models_ML_0_AR/y_test_NO2_201701_201709.rds")

y_validation_NO2_201710_201712 <- validation_NO2_201710_201712 %>% select(NO2_0)
X_validation_NO2_201710_201712 <- validation_NO2_201710_201712 %>% select(-NO2_0)

write_csv(X_validation_NO2_201710_201712, "train_test/Prediction_Models_ML_0_AR/X_validation_NO2_201710_201712.csv")
write_csv(y_validation_NO2_201710_201712, "train_test/Prediction_Models_ML_0_AR/y_validation_NO2_201710_201712.csv")

saveRDS(X_validation_NO2_201710_201712, "train_test/Prediction_Models_ML_0_AR/X_validation_NO2_201710_201712.rds")
saveRDS(y_validation_NO2_201710_201712, "train_test/Prediction_Models_ML_0_AR/y_validation_NO2_201710_201712.rds")

# We remove all objects from the workspace

rm(list = ls())

```

