---
title: "Train / test datasets for Python Models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading packages
```{r , warning= FALSE, message= FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(stringr)
library(knitr)
library(ranger)
library(broom)
library(RcppRoll)
```

As we are going to start to work with auto regressive models, we are going to need dataframes with the values of the variable to forecast (the target) and lagged versions of itself as inputs. 

## PM10 datasets auto regressive models

First we create the initial datasets for training, testing and validation for the PM10 variable. 
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

# We load the constitucion_data dataset
constitucion_data <- readRDS("data_rds/constitucion_data.rds") 

# We extract the periods of data that we are going to use 

train_PM10_201401_201612 <- constitucion_data %>% filter(date_time_utc >= '2014-01-01 00:00:00',
                                       date_time_utc <= '2016-12-31 23:00:00') 

# We generate a smaller training dataset with the last three months of 2016
train_PM10_201610_201612 <- constitucion_data %>% filter(date_time_utc >= '2016-10-01 00:00:00',
                                       date_time_utc <= '2016-12-31 23:00:00') 

# We generate a bigger training dataset with seven years of data
train_PM10_200901_201612 <- constitucion_data %>% filter(date_time_utc >= '2009-01-01 00:00:00',
                                       date_time_utc <= '2016-12-31 23:00:00') 
                                       

test_PM10_201701_201709 <- constitucion_data %>% filter(date_time_utc >= '2017-01-01 00:00:00',
                                       date_time_utc <= '2017-09-30 23:00:00') 

# We generate a smaller testing dataset with the first two weeks of 2016. 
test_PM10_20170101_20170114 <- constitucion_data %>% filter(date_time_utc >= '2017-01-01 00:00:00',
                                       date_time_utc <= '2017-01-14 23:00:00') 

validation_PM10_201710_201712 <- constitucion_data %>% filter(date_time_utc >= '2017-10-01 00:00:00',
                                       date_time_utc <= '2017-12-31 23:00:00') 

# We include all the datasets in a list
datasets_list <- list(train_PM10_201401_201612, train_PM10_201610_201612, train_PM10_200901_201612, test_PM10_201701_201709, test_PM10_20170101_20170114, validation_PM10_201710_201712)

# And we apply the same transformations to every dataframe in the list
datasets_list <- datasets_list %>% 
            map(~ select(., PM10)) %>% # We select the PM10 variable.
            map(~ mutate(., PM10 = replace_na(PM10, mean(PM10, na.rm = TRUE)))) %>% # We replace the NAs by the mean.
            map(~ rename(., PM10_0 = PM10)) # And we change the name of the variable to "PM10_0"


```

We add to the dataframes the PM10 lagged variables.
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

datasets_list <- datasets_list %>%
                  map(~ mutate(., PM10_1 = lag(PM10_0, 1),
                  PM10_2 = lag(PM10_0, 2),
                  PM10_3 = lag(PM10_0, 3),
                  PM10_4 = lag(PM10_0, 4),
                  PM10_5 = lag(PM10_0, 5),
                  PM10_6 = lag(PM10_0, 6),
                  PM10_7 = lag(PM10_0, 7),
                  PM10_8 = lag(PM10_0, 8),
                  PM10_9 = lag(PM10_0, 9),
                  PM10_10 = lag(PM10_0, 10),
                  PM10_11 = lag(PM10_0, 11),
                  PM10_12 = lag(PM10_0, 12),
                  PM10_13 = lag(PM10_0, 13),
                  PM10_14 = lag(PM10_0, 14),
                  PM10_15 = lag(PM10_0, 15),
                  PM10_16 = lag(PM10_0, 16),
                  PM10_17 = lag(PM10_0, 17),
                  PM10_18 = lag(PM10_0, 18),
                  PM10_19 = lag(PM10_0, 19),
                  PM10_20 = lag(PM10_0, 20),
                  PM10_21 = lag(PM10_0, 21),
                  PM10_22 = lag(PM10_0, 22),
                  PM10_23 = lag(PM10_0, 23),
                  PM10_24 = lag(PM10_0, 24),
                  PM10_25 = lag(PM10_0, 25),
                  PM10_26 = lag(PM10_0, 26),
                  PM10_27 = lag(PM10_0, 27),
                  PM10_48 = lag(PM10_0, 48),
                  PM10_72 = lag(PM10_0, 72),
                  PM10_96 = lag(PM10_0, 96),
                  PM10_120 = lag(PM10_0, 120),
                  PM10_144 = lag(PM10_0, 144),
                  PM10_168 = lag(PM10_0, 168))) 


# We unlist the dataframes and we rename them again.
for (i in seq(datasets_list))
  assign(paste0("df", i), datasets_list[[i]])

train_PM10_201401_201612 <- df1 %>% na.omit()
train_PM10_201610_201612 <- df2 %>% na.omit()
train_PM10_200901_201612 <- df3 %>% na.omit()
test_PM10_201701_201709 <- df4 %>% na.omit()
test_PM10_20170101_20170114 <- df5 %>% na.omit()
validation_PM10_201710_201712 <- df6 %>% na.omit()

```


We create and save all the y, X datasets as rds files 
```{r warning = FALSE, message= FALSE, out.width= '\\textwidth'}

y_train_PM10_201401_201612 <- train_PM10_201401_201612 %>% select(PM10_0)
X_train_PM10_201401_201612 <- train_PM10_201401_201612 %>% select(-PM10_0)

saveRDS(X_train_PM10_201401_201612, "train_test/Forecasting_Models_ML_PM10_AR/X_train_PM10_201401_201612.rds")
saveRDS(y_train_PM10_201401_201612, "train_test/Forecasting_Models_ML_PM10_AR/y_train_PM10_201401_201612.rds")


y_train_PM10_201610_201612 <- train_PM10_201610_201612 %>% select(PM10_0)
X_train_PM10_201610_201612 <- train_PM10_201610_201612 %>% select(-PM10_0)


saveRDS(X_train_PM10_201610_201612, "train_test/Forecasting_Models_ML_PM10_AR/X_train_PM10_201610_201612.rds")
saveRDS(y_train_PM10_201610_201612, "train_test/Forecasting_Models_ML_PM10_AR/y_train_PM10_201610_201612.rds")

y_train_PM10_200901_201612 <- train_PM10_200901_201612 %>% select(PM10_0)
X_train_PM10_200901_201612 <- train_PM10_200901_201612 %>% select(-PM10_0)


saveRDS(X_train_PM10_200901_201612, "train_test/Forecasting_Models_ML_PM10_AR/X_train_PM10_200901_201612.rds")
saveRDS(y_train_PM10_200901_201612, "train_test/Forecasting_Models_ML_PM10_AR/y_train_PM10_200901_201612.rds")


y_test_PM10_20170101_20170114 <- test_PM10_20170101_20170114 %>% select(PM10_0)
X_test_PM10_20170101_20170114 <- test_PM10_20170101_20170114 %>% select(-PM10_0)


saveRDS(X_test_PM10_20170101_20170114, "train_test/Forecasting_Models_ML_PM10_AR/X_test_PM10_20170101_20170114.rds")
saveRDS(y_test_PM10_20170101_20170114, "train_test/Forecasting_Models_ML_PM10_AR/y_test_PM10_20170101_20170114.rds")

y_test_PM10_201701_201709 <- test_PM10_201701_201709 %>% select(PM10_0)
X_test_PM10_201701_201709 <- test_PM10_201701_201709 %>% select(-PM10_0)


saveRDS(X_test_PM10_201701_201709, "train_test/Forecasting_Models_ML_PM10_AR/X_test_PM10_201701_201709.rds")
saveRDS(y_test_PM10_201701_201709, "train_test/Forecasting_Models_ML_PM10_AR/y_test_PM10_201701_201709.rds")

y_validation_PM10_201710_201712 <- validation_PM10_201710_201712 %>% select(PM10_0)
X_validation_PM10_201710_201712 <- validation_PM10_201710_201712 %>% select(-PM10_0)

saveRDS(X_validation_PM10_201710_201712, "train_test/Forecasting_Models_ML_PM10_AR/X_validation_PM10_201710_201712.rds")
saveRDS(y_validation_PM10_201710_201712, "train_test/Forecasting_Models_ML_PM10_AR/y_validation_PM10_201710_201712.rds")

# We remove all objects from the workspace

rm(list = ls())

```

