---
title: "Gijon_Air_Pollution"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gijon Air Pollution notes

Notes about the development of my capstone final project of my Master on Data Science in KSchool.

### Project goals:

  - Construct a web dashboard to visualize the evolution of the main air pollution indicators of Gijon city. (Asturias; Spain).
  - Build a model to try to predict the probability of exceeding the pollution limits established by the EU in the city of Gijon (one day in advance).

#### History

2018-08-16 
I created my github repository.
I set the main goals of the project.

#### References

- OpenAir package web: http://davidcarslaw.github.io/openair/
- Breezometer: https://breezometer.com/real-time-air-quality

#### Data gathering

Loading packages
```{r }
library(readr)
library(dplyr)
library(openair) # http://davidcarslaw.github.io/openair/

```


First of all I have to check if I will have the basic data to make the analysis. I need air pollution and weather data of the Gijon area. 
The town hall of Gijon has an open data web portal here https://transparencia.gijon.es/. We can download pollution air data on several formats from year 2000 to 2017 here: https://transparencia.gijon.es/search/risp_dataset/page/1808-catalogo-de-datos?utf8=%E2%9C%93&search=aire+&search_sector=&search_format=&commit=Buscar&authenticity_token=j8%2F3CvCuPcDkrRe%2F1NR5RBp0t%2FOOosiA7724w3T2mB4%3D

I download the data of the year 2017 with csv format.

I import the csv file to R in order to give a first lood to the data

```{r }
air_data_2017 <- read_csv('data/air_data_2017.csv', locale = locale(encoding = "ISO-8859-1"))
head(air_data_2017)
```


```{r }
summary(air_data_2017)
```


